\chapter{Gauss Norm}

Thus far we have shown multivariate restricted power series with paramter $c$ over a non-archimedean
normed ring $R$ in $n$ variables form a ring. Let us again denote this ring by
$R_c [ \! [ x_1, \dots, x_n ] \! ]$.

\begin{definition}
Consider the following function $\lvert \cdot \rvert _c :  R_c [ \! [ x_1, \dots, x_n ] \! ] \to \R$
defined by
\[
    \lvert f \rvert _c = \sup_{i \in \mathbb{N}^n} \lvert a_{(i_1, \cdots, i_n)} \rvert
    \prod_n \lvert c \rvert _n^{i_n}.
\]
for $f(x) = \sum a_{(i_1, \cdots, i_n)} x_1^{i_1} \cdots x_n^{i_n}$ and call this the Gauss norm
with parameter $c$.
\end{definition}

We want to prove that this is in fact a norm and is non-archimedean.

\begin{lemma}
  Let $c$ be strictly positive, that is for all $i, 0 < c i$, then the pair
  $( R_c [ \! [ x_1, \dots, x_n ] \! ], | \cdot |_c)$ is a non-archimedean normed ring.
\end{lemma}

\section{The proof}

We break this statement up into small lemmas that together prove our goal.

\begin{lemma}
  \[
    \lvert 0 \rvert _c = 0.
  \]
\end{lemma}

\begin{proof}
  Immediate.
\end{proof}

\begin{lemma}
  Gauss norm is non-negative.
\end{lemma}

\begin{proof}
  Consider when $i = (0, \cdots, 0) = 0$, then $\lvert a_0 \rvert \leq \lvert f \rvert _c.$
  We have the LHS is greater than or equal to 0 since $R$ is a normed ring. So we are done.
\end{proof}

\begin{lemma}
  If $c$ is strictly positive, then
  \[
    \lvert f \rvert _c 0 \iff f = 0.
  \]
\end{lemma}

\begin{proof}
  We only need the only if part. Prove this by contradiction, that is want to show $f \neq 0
    \implies \lvert f \rvert _c \neq 0.$ This is immediate from noticing $f$ has a non-zero
    coeffecient and since $c$ strictly positive, evaluating at the monomial with non-zero coeff, we
    get a positive value of $\lvert a_{(i_1, \cdots, i_n)} \rvert \prod_n \lvert c \rvert _n^{i_n}$
    and so we are done.
\end{proof}

\begin{lemma}
  Gauss norm is non-archimedean.
\end{lemma}

\begin{proof}

\end{proof}

\section{Polynomials}

Moreover, if we restrict ourselves to polynomials we can strenghten this.

\begin{lemma}
  The pair $(R[x], | \cdot |_c)$ is a non-archimedean valued ring. That is, $\lvert \cdot \rvert _c$
  is an absolute value.
\end{lemma}

%% Note BGR actually shows that it is a an absolute value -- so probably should just show that first
