\chapter{Restricted Power Series}

\section{Definition}

Recall formal multivariate power series over a set $K$ in $n$ variables, written
$K [ \! [ x_1, \dots, x_n ] \! ]$ are infinite sums
\[
  f(x) = \sum a_{(i_1, \cdots, i_n)} x_1^{i_1} \cdots x_n^{i_n},
\]
where the sum is taken over all $n$-tuples $(i_1, \cdots i_n)$ with $i_j \in \mathbb{N}$ and
$a_{(i_1, \cdots, i_n)} \in K$.

When $K$ is a ring, formal multivariate power series in $n$ variables form a ring, and we will refer
to them as the ring of power series over $K$ in $n$ variables. We again denote such
rings using $K [ \! [ x_1, \dots, x_n ] \! ]$.

We define a subset of $K [ \! [ x_1, \dots, x_n ] \! ]$ that consists of power series whose
coefficients approach zero in the cofinite filter.

\begin{definition}
  Let $f$ be a power series over a normed ring $R$ in $n$ variables, write
  $f(x) = \sum a_{(i_1, \cdots, i_n)} x_1^{i_1} \cdots x_n^{i_n}$, and let
  $c = (c_1, \cdots, c_n)$ be an $n$-tuple of real numbers.
  We say $f$ is a restricted power series of parameter $c$ over $R$ in $n$ variables if
  \[
    \lim_{n \to \infty} \lvert a_{(i_1, \cdots, i_n)} \rvert \prod_{j=1}^{n} c_j^{i_j} = 0
  \]
  in the cofinite topology.

  We denote the set of restricted power series of parameter $c$ over $R$ in $n$ variables by
  $R_c \langle \! [ x_1, \dots, x_n ] \! \rangle$.
\end{definition}

\begin{remark}
  Note
  \[
    \lim \lvert a_{(i_1, \cdots, i_n)} \rvert \prod_{j=1}^{n} c_j^{i_j} = 0
  \]
  in the cofinite topology, if for all $0 < \epsilon$ there are finitely many $n$-tuples that have
  $\left \lvert \lvert a_{(i_1, \cdots, i_n)} \rvert \prod_{j=1}^{n} c_j^{i_j} \right \rvert$
  greater than $\varepsilon$.
\end{remark}

Moreover, it actually does not matter about the sign of the values $c_i$, thus we may as well consider
$c_i > 0$ for all $i$.

\begin{lemma}
  A power series $f$ is restricted for parameter $c$ if and only if it is restricted for parameter
  $\lvert c \rvert$.
\end{lemma}

\section{Additive Group}

It turns out that restricted power series of parameter $c$ over a normed ring $R$ in $n$ variables
are not just a set, in fact they form an additive group.

\begin{lemma}
  For any normed ring $R$, natural number $n$, and tuple $c$,
  $R_c \langle \! [ x_1, \dots, x_n ] \! \rangle$ is an additive group.
\end{lemma}

\section{Ring}

Moreover, if we strengthen our condition of normed ring $R$, to a non-archimedean normed ring, we
in fact can show it is a ring.

\begin{theorem}
  For any non-archimedean normed ring $R$, $n \in \mathbb{N}$ and $n$-tuple $c \in \mathbb{R}^n$,
  $R_c \langle \! [ x_1, \dots, x_n ] \! \rangle$ is a ring.
\end{theorem}

The only extra thing we need to show is closure of multiplication. That is the product of two restricted
power series are also restricted.

To prove this in Lean, we instead give a more general statement that will give what we want as an
immediate consequence.

\begin{lemma}
  Let $M$ be an additive monoid with an antidiagonal, and $S$ a non-archimedean normed ring. Let
  $\phi : M \to \mathbb{R}$ be a function such that for all $a, b \in M, \phi(a + b) = \phi(a)
  \phi(b)$. If two functions $f, g : M \to S$ have the property that
  $\lim_i \lvert f i \rvert \phi (i) = 0$ in the cofinite filter and
  $\lim_i \lvert g i \rvert \phi (i) = 0$ then
  \[
  \lim_i \lvert \sum_{\mathclap{\substack{(a,b) \in \\ \textrm{antidiagonal } i}}} f(a) g(b) \rvert
  \phi(i) = 0.
  \]
\end{lemma}

\section{Tate algebra}

With this new definition we can now define a necessary component of our project.

\begin{definition}
  When $c \equiv 1$ and $R$ is a complete non-archimedean field, we call the ring of restricted power
  series over $R$ in $n$ variables, the \emph{Tate algebra} over $K$ in $n$ variables, we denote this
  by $T_n$.
\end{definition}

We can consider Tate algebras not just with variable $c \equiv 1$, that is more general polydiscs,
these turn out to be examples of affiniod algebras. See the Affinoid chapter for more information.


%% Note the following really needs to go below gauss norm as we need the ring of restricted power series
%% to be a normed ring (with the gauss norm!)

\section{Isomorphism}

When considering multivariate power series, it is convenient to consider the following "equality"
\[
  K [ \! [ x_1, \dots, x_n ] \! ] = (K [ \! [ x_1, \dots, x_{n-1} ] \! ])[\![ x_n ]\!]
\]
this is seen by writing $f \in K [ \! [ x_1, \dots, x_n ] \! ]$ in the form
\[
  f = \sum_{k = 0}^\infty g_k (x_1, \dots, x_n) x_n^k.
\]
Of course, by "equality" we really mean the spaces are isomorphic. We would like to have a similar
isomorphism for multivariate restricted power series.

\begin{lemma}
  $R_c \langle x_1, \dots, x_{n + 1} \rangle \equiv
  (R_c \langle x_1, \dots, x_n \rangle) \langle x_{n+1} \rangle.$
\end{lemma}

\begin{proof}
Let $\phi$ be the map given by the isomorphism of multivariate power series detailed above. It
suffices to show that the image by $phi, \phi^{-1}$ of restricted power series are restricted.

Assume for all $1 \leq m \leq n+1$, $c_m \neq 0$, else every power series is restricted trivially
and there is nothing to prove.

Towards this let $f = \sum a_{i_1, \dots i_{n+1}} x_1^{i_1} \cdots x_{n+1}^{i+1} \in
R_c \langle x_1, \dots, x_{n + 1} \rangle$. It being restricted means
\[
  \lim \lvert a_{(i_0, \cdots, i_{n+1})} \rvert \lvert c \rvert^{(i_0, \cdots, i_{n+1})} = 0
\]
under the cofinite filter. Write
\[
  \phi (f) = \sum_{s = 0}^\infty g_s x_n^s
\]
where $g_s \in R [ \! [ x_1, \dots, x_n ] \! ]$.
It suffices to show two things:
\begin{enumerate}
  \item for all $s$, $g_s \in R_c \langle x_1, \dots, x_n \rangle$; and
  \item $\phi (f)$ is restricted in $(R \langle x_1, \dots, x_{n-1} \rangle)[ \! [x_n] \! ]$.
\end{enumerate}

For 1. note for a fixed $s$, $g_s = \sum a_{((i_0, \cdots, i_n),s)} x_1^{i_1} \cdots x_n^{i_n}$;
thus we want to show
\[
  \lim \lvert a_{((i_0, \cdots, i_n), s)} \rvert \prod_{j=1}^{n} \lvert c_j \rvert ^{i_j} = 0
\]
under the cofinite filter. Suppose it does not, that is there exists an $\varepsilon$ such that there
are infinitely many $(i_0, \cdots, i_n)$ with
\[
  \lvert a_{((i_0, \cdots, i_n), s)} \rvert \prod_{j=1}^{n} \lvert c_j \rvert ^{i_j} > \varepsilon.
\]
But then for this fixed $s$ there are infinitely many $(i_0, \cdots, i_n)$ with
\[
  \left( \lvert a_{((i_0, \cdots, i_n), s)} \rvert \prod_{j=1}^{n} \lvert c_j \rvert ^{i_j} \right)
  \lvert c_n \rvert^s > \varepsilon \lvert c_n \rvert^s
\]
which would contradicted the restricted-ness of $f$ - because we have infinitely many $(i_1, \cdots, i_{n+1})$,
namely $(i_1, \cdots, i_n, s)$ with
\[
  \lvert a_{(i_0, \cdots, i_{n+1})} \rvert \lvert c \rvert^{(i_0, \cdots, i_{n+1})} > \varepsilon' =
  \varepsilon \lvert c_n \rvert^s.
\]
Since this is independent of the choice of $s$, we have shown as wanted.

For 2. we want to show
\[
  \lim \lvert g_s \rvert \lvert c_n \rvert^s = 0
\]
in the cofinite filter. Recall by the definition of the gauss norm,
\[
  \lvert g_s \rvert = \max \lvert a_{(i_0, \cdots, i_n, s)} \rvert \prod_{j=1}^{n} \lvert c_j \rvert ^{i_j}
\]
and moreover, since $g_s$ is a restricted power series the gauss norm is attained. That is
there exists $(k_{s_1}, \cdots k_{s_n})$ such that
\[
  \lvert g_s \rvert = \lvert a_{(k_{s_1}, \cdots k_{s_n}, s)} \rvert \prod_{j=1}^{n} \lvert c_j \rvert ^{i_j}.
\]
Now suppose $\phi (f)$ is not restricted, that is there exists an $\varepsilon$ such that there are
infinitely many $s$ with
\[
  \left( \lvert a_{((k_{s_1}, \cdots k_{s_n}), s)} \rvert \prod_{j=1}^{n} \lvert c_j \rvert ^{i_j} \right)
  \lvert c_n \rvert^s > \varepsilon.
\]
But once again this implies there are infinitely many $(i_1, \cdots, i_{n+1})$ such that
\[
  \lvert a_{(i_0, \cdots, i_{n + 1})} \rvert \prod_{j=1}^{n} \lvert c_j \rvert ^{i_j} > \varepsilon,
\]
contradicting $f$ being restricted.

For the other direction consider $p = \sum_{s = 0}^\infty q_s x_n^s \in
(R_c \langle x_1, \dots, x_n \rangle) \langle x_{n+1} \rangle$, with
$q_s = \sum a_{(i_{s_1}, \cdots, i_{s_n})} x_1^{i_{s_1}} \cdots x_n^{i_{s_n}}.$ Then
\[
  \phi^{-1} (p) = \sum b_{(i_1, \cdots, i_{n+1})} x_1^{i_1} \cdots x_n^{i_{n+1}},
\]
where $b_{(i_1, \cdots, i_n, s)} = a_{(i_{s_1}, \cdots, i_{s_n})}$.
Suppose $\varphi(p)$ is not restricted. Then there exists an $\varepsilon$ such that there are
infinitely many tuples $(i_1, \cdots, i_{n+1})$ with
\[
  \lvert b_{(i_1, \cdots, i_{n+1})} \rvert \prod_{j=1}^{n+1} \lvert c_j \rvert ^{i_j} > \varepsilon.
\]

Suppose in this infinite set that $i_{n+1}$ only takes finitely many values. Then for some $s$,
there are infinitely many $(i_{s_1}, \cdots, i_{s_n})$ such that $(i_{s_1}, \cdots, i_{s_n}, s)$
belongs in the set. However, this means there are infinitely many $(i_{s_1}, \cdots, i_{s_n})$ such that
\[
  \left( \lvert  b_{(i_{s_1}, \cdots, i_{s_n}, s)} \rvert \prod_{j=1}^{n} \lvert c_j \rvert ^{i_{s_j}} \right) \lvert c_n \rvert^s =
  \left( \lvert  a_{(i_{s_1}, \cdots, i_{s_n})} \rvert \prod_{j=1}^{n} \lvert c_j \rvert ^{i_{s_j}} \right) \lvert c_n \rvert^s > \varepsilon.
\]
Thus, setting $\varepsilon' = \varepsilon / \lvert c_n \rvert^s$ we have infinitely many
$(i_1, \cdots, i_n)$ such that the following inequality holds
\[
  \lvert  a_{(i_{s_1}, \cdots, i_{s_n})} \rvert \prod_{j=1}^{n} \lvert c_j \rvert ^{i_j} > \varepsilon'
\]
which contradicts the restrictedness of $q_s$.

On the other hand suppose $i_{n+1}$ takes infinitely many values, that is there are infinitely many
$s$ such that
\[
  \left( \lvert  b_{(i_{s_1}, \cdots, i_{s_n}, s)} \rvert \prod_{j=1}^{n} \lvert c_j \rvert ^{i_{s_j}} \right) \lvert c_n \rvert^s
  > \varepsilon.
\]
But after relabeling
\[
  \left( \lvert  b_{(i_{s_1}, \cdots, i_{s_n}, s)} \rvert \prod_{j=1}^{n} \lvert c_j \rvert ^{i_{s_j}} \right) \lvert c_n \rvert^s  =
  \lvert  a_{(k_1, \cdots, k_n)} \rvert \prod_{j=1}^{n+1} \lvert c_j \rvert ^{k_j}.
\]
Thus, for the infinitely many $s$ in the set
\begin{align*}
  \lvert g_s \rvert \lvert c_n \rvert^s & = \max \left\{ \lvert  a_{(i_{s_1}, \cdots, i_{s_n})} \rvert
  \prod_{j=1}^{n+1} \lvert c_j \rvert ^{k_j} \right\} \lvert c_n \rvert ^s \\
  & \geq \lvert  a_{(k_1, \cdots, k_n)} \rvert \prod_{j=1}^{n+1} \lvert c_j \rvert ^{k_j} \\
  & > \varepsilon
\end{align*}
contradicting the restrictedness of $p$.
\end{proof}
